Manuscript 1 code

### 2024-10-11 manuscript_1_code.Rmd
Added option for alpha and gamma stability

@@

## set up site-level data for model
```{r site cover by diversity data setup}

@@

# calculate mean synchrony within a site (!= synchrony of species at the site level!)
mean_synchrony <- diversity_stability_synchrony %>% 
  group_by(habitat, site) %>% 
  summarise(synchrony_mean = mean(synchrony, na.rm = T))

# merge with site level dss metrics
dss_spatial <- merge(diversity_stability_site, spatial_synchrony, by = "site_habitat")
dss_spatial_2 <- merge(dss_spatial, mean_synchrony, by = c("habitat", "site"))

dss_spatial_ranges <- dss_spatial_2 %>%
  group_by(habitat) %>%
  summarise(min_spatial_synchrony = min(spatial_synchrony, na.rm = TRUE),
            max_spatial_synchrony = max(spatial_synchrony, na.rm = TRUE),
            min_mean_synchrony= min(synchrony_mean, na.rm = TRUE),
            max_mean_synchrony = max(synchrony_mean, na.rm  = TRUE))

@@

## panel graph
```{r figure 5 panel, fig.height = 6, fig.width=22}
(figure_5 <- ggarrange(rich_synch_plot, figure_5b, figure_5a, 
                      ncol = 3, nrow = 1, 
                      common.legend = TRUE,
                      legend = "bottom", 
                      labels = c("a.", "b.", "c."),
                      font.label = list(size = 26, color = "black", face = "plain")))

# ggsave(filename = "output/figure_5_option2.png", figure_5, height = 8, width = 16)
```

>>>>

```{r add cover to site df}

@@

@@

# calculate mean synchrony within a site (!= synchrony of species at the site level!)
mean_synchrony_stability <- diversity_stability_synchrony %>% 
  group_by(habitat, site) %>% 
  summarise(synchrony_mean = mean(synchrony, na.rm = T),
            stability_mean = mean(cover_stability, na.rm = T))

# merge with site level dss metrics
dss_spatial <- merge(diversity_stability_site, spatial_synchrony, by = "site_habitat")
dss_spatial_2 <- merge(dss_spatial, mean_synchrony_stability, by = c("habitat", "site"))

dss_spatial_ranges <- dss_spatial_2 %>%
  group_by(habitat) %>%
  summarise(min_spatial_synchrony = min(spatial_synchrony, na.rm = TRUE),
            max_spatial_synchrony = max(spatial_synchrony, na.rm = TRUE),
            min_mean_synchrony= min(synchrony_mean, na.rm = TRUE),
            max_mean_synchrony = max(synchrony_mean, na.rm  = TRUE),
            min_mean_stability = min(stability_mean, na.rm = TRUE),
            max_mean_stability = max(stability_mean, na.rm  = TRUE))

@@

```{r alpha gamma stability mod plot}
alpha_gamma_stab_mod <- glmmTMB(cover_stability ~ stability_mean*habitat, family = Gamma("log"), data = dss_spatial_2)
summary(alpha_gamma_stab_mod)
Anova(alpha_gamma_stab_mod)
#                          Chisq Df Pr(>Chisq)    
# stability_mean         61.2725  1   4.97e-15 ***
# habitat                12.9323  3   0.004785 ** 
# stability_mean:habitat  9.4832  3   0.023511 *

emtrends(alpha_gamma_stab_mod, pairwise ~ habitat, var = "stability_mean") # backreef different from fringe and forereef 10
# Fringing - Backreef            1.078 0.381 Inf   2.828  0.0242
# Fringing - Forereef 10m        0.175 0.345 Inf   0.509  0.9571
# Fringing - Forereef 17m        0.380 0.305 Inf   1.248  0.5965
# Backreef - Forereef 10m       -0.903 0.349 Inf  -2.589  0.0475
# Backreef - Forereef 17m       -0.698 0.309 Inf  -2.256  0.1085
# Forereef 10m - Forereef 17m    0.205 0.263 Inf   0.779  0.8638

alpha_gamma_emm <- emmip(alpha_gamma_stab_mod,
                        habitat ~ stability_mean,
                        at = list(stability_mean = seq(0,10,0.01)), plotit = F, CIs = T)

filtered_ag_effects <- alpha_gamma_emm %>%
  left_join(dss_spatial_ranges, by = c("habitat")) %>%
  filter(stability_mean >= min_mean_stability & 
           stability_mean <= max_mean_stability)

(alpha_gamma_stability_plot <- 
  ggplot() +
  geom_point(data = dss_spatial_2,
             aes(x = stability_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_ag_effects, aes(x = stability_mean, y = exp(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_ag_effects, aes(x = stability_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Site stability\n", x = "Plot stability", title = "") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=24), 
        axis.title.y = element_text(size=24), 
        axis.text.x = element_text(size=22), 
        axis.text.y = element_text(size=22),
        legend.text = element_text(size=22),
        plot.title = element_text(size=26), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )
```

## panel graph
```{r figure 5 panel, fig.height = 6, fig.width=22}
(figure_5 <- ggarrange(figure_5b, figure_5a, alpha_gamma_stability_plot, 
                      ncol = 3, nrow = 1, 
                      common.legend = TRUE,
                      legend = "bottom", 
                      labels = c("a.", "b.", "c."),
                      font.label = list(size = 26, color = "black", face = "plain")))

# ggsave(filename = "output/figure_5_option3.png", figure_5, height = 8, width = 16)
```
